### 2014-06-01 ACL2014/system.tex
section 3 edits

\subsection{\lsh for Cosine Similarity}
\label{subsec:vlsh:cosine}
In this paper, we are interested in cosine similarity, and use the \lsh family defined by Charikar \shortcite{Charikar02STOC}. 
For two queries; $p,q \in \mathbb{R}^D$, the cosine similarity between them is $\left( \frac{p.q}{ \| p \| \| q \| }\right)$. 
The \lsh functions for cosine similarity is defined as follows: if $\alpha \in \mathbb{R}^D$ is a random vector, 
then a corresponding hash function $h_{\alpha}$ can be defined as $h_{\alpha}(p) = \mathrm{sign}(\alpha \cdot p)$.
Typically, a negative sign is represented as $0$ and positive sign as $1$, and indices of buckets in the hash tables (i.e. the
range of each $g_j$) are $K$ bit vectors. 
In order to create a random vector $\alpha$, we exploit the intuition in
\cite{Achlioptas03,LiH06} and sample each coordinate of $\alpha$ from $\{-1, +1\}$ with equal probability.
Practically, each coordinate of $\alpha$ is generated using a hash function that maps that coordinate
to $\{-1, +1\}$ (this is termed ``hashing trick'' in~\cite{}). 
%\footnote{Note, we represent our queries ($p$ and $q$) using only non-zero features.}  
This hashing trick is important and useful as we do not need to explicitly store the 
huge random projection matrix of size $D \times K \times L$.

%To solve the above problem, we propose a distributed Hadoop framework based on Locality Sensitive Hashing. 
%Our framework has three main steps
%i) Generate $256$ bit signature for each query.  
%ii) Use the $256$ bit signature for each query to generate $L$ tables for each table with $K$ bit signature. 
%Next, generate query pairs which fall in the same table.
%iii) Compute cosine similarity between query terms which fall in the same table and return the query pairs  
%with similarity $\geq\tau$.  
%
Figure~\ref{algo:lsh} describes the algorithm for both creating the data structure, as well as for
querying it. In preprocessing step, the algorithm takes as input $N$ queries along with the associated feature vectors. 
In our setting, each query is represented using an extremely sparse and high dimensional feature vector
that is constructed as follows: for query $q$, we take all the webpages (urls) that any user has clicked on 
when querying for $q$. Using this representation, we then generate the 
$L$ different hash values for each query $q$, where each such hash value is again 
the concatenation of $K$ hash functions. These $L$ hash values per query are then used to create $L$ hash tables.
Since the width of the index of each bucket $K$, each hash table contains at most $2^K$ buckets. 
with each hash table containing at most $2^K$ buckets.
Each query term is then placed in their respective buckets for each of the $L$ hash tables.

In querying step, for input $M$ test queries, we retrieve all the query 
terms which appear in the buckets associated with $M$ test queries. 
Next, compute cosine similarity between each of the retrieved terms 
and the input test queries. Return all those queries as neighbors 
which are within a similarity threshold ($\tau$).

In this research, we have implemented the above algorithm within the distributed setting (Hadoop).   
In our experiments, we show that the algorithm scales to hundreds of millions of queries (see Section \ref{subsec:eval:vanillaLSH}).

\begin{figure}
\fbox{
\begin{minipage}{2.5in}
{\bf Preprocessing:} Input is $N$ queries with their respective feature vectors 
\begin{itemize}
\item Select $L$ functions $g_j$, $j=1,2,\dots,L$, setting  $g_j(q)=(h_{1,j}(q),h_{2,j}(q),\dots,h_{K,j}(q)))$, where $h_{1,j}(q),h_{2,j}(q),\dots,h_{K,j}(q)))$ are chosen at random from the \lsh family.
\item Construct $L$ hash tables, $\forall 1 \leq j \leq  L$. All queries with similar $g_j$ functions ($\forall 1 \leq j \leq  L$) are placed in the same bucket.    
\end{itemize}

{\bf Query:} Input is queries ($q=1,2,\dots,M$); $M$ is the number of test queries
\begin{itemize}
\item For each $j=1,2,\dots,L$
\begin{itemize}
\item Retrieve all the queries from the bucket with $g_j(q)$ function as the index of the bucket 
\item Compute cosine similarity between query q and all the retrieved queries. Return all the queries which have similarity affinity within a similarity threshold ($\tau$). 
\end{itemize}

\end{itemize}
\end{minipage}
}
\caption{\lshf Algorithm}
\end{figure}

>>>>

\subsection{\lsh for Cosine Similarity}
\label{subsec:vlsh:cosine}
In this paper, we are interested in cosine similarity, and use the \lsh family defined by Charikar \shortcite{Charikar02STOC}. 
For two queries; $p,q \in \mathbb{R}^D$, the cosine similarity between them is $\left( \frac{p.q}{ \| p \| \| q \| }\right)$. 
The \lsh functions for cosine similarity is defined as follows: if $\alpha \in \mathbb{R}^D$ is a random vector, 
then a corresponding hash function $h_{\alpha}$ can be defined as $h_{\alpha}(p) = \mathrm{sign}(\alpha \cdot p)$.
Typically, a negative sign is represented as $0$ and positive sign as $1$, and indices of buckets in the hash tables (i.e. the
range of each $g_j$) are $K$ bit vectors. 
In order to create a random vector $\alpha$, we exploit the intuition in
\cite{Achlioptas03,LiH06} and sample each coordinate of $\alpha$ from $\{-1, +1\}$ with equal probability.
Practically, each coordinate of $\alpha$ is generated using a hash function that maps that coordinate
to $\{-1, +1\}$ (this is termed ``hashing trick'' in~\cite{weinberger09hashing}). 
%\footnote{Note, we represent our queries ($p$ and $q$) using only non-zero features.}  
This hashing trick is important and useful as we do not need to explicitly store the 
huge random projection matrix of size $D \times K \times L$.

%To solve the above problem, we propose a distributed Hadoop framework based on Locality Sensitive Hashing. 
%Our framework has three main steps
%i) Generate $256$ bit signature for each query.  
%ii) Use the $256$ bit signature for each query to generate $L$ tables for each table with $K$ bit signature. 
%Next, generate query pairs which fall in the same table.
%iii) Compute cosine similarity between query terms which fall in the same table and return the query pairs  
%with similarity $\geq\tau$.  
%
Figure~\ref{alg:lsh} describes the algorithm for both creating the data structure, as well as for
querying it. In preprocessing step, the algorithm takes as input $N$ queries along with the associated feature vectors. 
In our setting, each query is represented using an extremely sparse and high dimensional feature vector
that is constructed as follows: for query $q$, we take all the webpages (urls) that any user has clicked on 
when querying for $q$. Using this representation, we then generate the 
$L$ different hash values for each query $q$, where each such hash value is again 
the concatenation of $K$ hash functions. These $L$ hash values per query are then used to create $L$ hash tables.
Since the width of the index of each bucket $K$, each hash table contains at most $2^K$ buckets. 
with each hash table containing at most $2^K$ buckets.
Each query term is then placed in their respective buckets for each of the $L$ hash tables.

In order to retrieve near neighbors, for each of the $M$ test queries, we first 
retrieve all the query terms which appear in the buckets associated with each of the $M$ test queries. 
Next, we compute cosine similarity between each of the retrieved terms 
and the input test queries and return all those queries as neighbors 
which are within a similarity threshold ($\tau$).

In this work, we have implemented the above algorithm in a map-reduce setting (Hadoop). 
In Section \ref{subsec:eval:vanillaLSH}, we show that the map-reduce implmentation scales to hundreds of millions of queries.

\begin{figure}
\fbox{
\begin{minipage}{2.5in}
{\bf Preprocessing:} Input is $N$ queries with their respective feature vectors.
\begin{itemize}
\item Select $L$ functions $g_j$, $j=1,2,\dots,L$, setting  $g_j(q)=(h_{1,j}(q),h_{2,j}(q),\dots,h_{K,j}(q))$, where 
$\{h_{i,j}(q), i\in [1, K], j\in [1,L]\}$ are chosen at random from the \lsh family.
\item Construct $L$ hash tables, $\forall 1 \leq j \leq  L$. 
All queries with similar $g_j$ functions ($\forall 1 \leq j \leq  L$) are placed in the same bucket.    
\end{itemize}

{\bf Query:} $M$ test queries. Let $q$ denote a test query. 
\begin{itemize}
\item For each $j=1,2,\dots,L$
\begin{itemize}
\item Retrieve all the queries from the bucket with $g_j(q)$ function as the index of the bucket 
\item Compute cosine similarity between query q and all the retrieved queries. Return all the queries which have similarity affinity within a similarity threshold ($\tau$). 
\end{itemize}

\end{itemize}
\end{minipage}
}
\caption{\lshf Algorithm}
\label{alg:lsh}
\end{figure}
